---
title: "HW1"
author: "Mariia Firuleva"
date: "22 10 2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F)
```


# Load libraries

```{r}
library(data.table)
library(ggplot2)
library(RColorBrewer)
library(dplyr)
library(tidyr)
library(ggpubr)
```


# 1

```{r}
df <- fread('insurance_cost.csv') %>% 
  mutate_if(is.character, as.factor)
knitr::kable(df %>% head())
```


```{r}
knitr::kable(table(df$children))
```

```{r}
df <- df %>% 
  mutate(children = factor(children))
```


# 2

```{r}
get_hist <- function(df, variable) {
  ggplot(df, aes_string(x = variable))+
    geom_histogram()+
    theme_bw()+
    theme(aspect.ratio = 1,
          plot.title = element_text(hjust = 0.5))+
    ggtitle(sprintf('%s', variable))
}

hists <- lapply(colnames(df %>% select_if(is.numeric)), function(col) get_hist(df, col))
ggarrange(plotlist = hists)
```


# 3

```{r}
p3 <- ggplot(data = df, aes(x = charges))+
  geom_density(alpha = 0.5)+
  geom_vline(xintercept = mean(df$charges), color='red')+
  geom_text(aes(x=mean(df$charges) + 1e3, label=sprintf("mean(charges) = %g", mean(df$charges)),
                y=0.00003), colour="red",
            angle=90, text=element_text(size=12)) +
  geom_vline(xintercept = median(df$charges), color='blue')+
  geom_text(aes(x=median(df$charges) - 1e3, label=sprintf("median(charges) = %g", median(df$charges)),
                y=0.00003), colour="blue",
            angle=90, text=element_text(size=12)) +
  theme_bw()+
  ylab('Density')+
  theme(aspect.ratio = 1)
p3
```


# 4

```{r, fig.width=15,fig.height=15}
get_bxplt <- function(df, x, y) {
  ggplot(data = df, aes_string(y = y, x = x, fill = x))+
    geom_boxplot()+
    theme_bw()+
    scale_fill_brewer(palette = 'Set1')+
    theme(aspect.ratio = 1)
}

bxplts <- lapply(c('sex', 'smoker', 'region'), function(x)
  get_bxplt(df = df, x = x, y = 'charges'))
p4 <- ggarrange(plotlist = bxplts, ncol = 3)
p4
```

# 5

```{r, fig.width=15,fig.height=15}
ggarrange(p3, p4, nrow = 2)
```


# 6

```{r}
mean_median_df <- df %>% 
  select(charges, region) %>% 
  group_by(region) %>% 
  summarise(mean=mean(charges), median=median(charges))


ggplot(df, aes(x = charges))+
  geom_density()+
  facet_wrap(~ region, scales='free_y')+
  geom_vline(data = mean_median_df, aes(xintercept = mean), color="red")+
  geom_vline(data = mean_median_df, aes(xintercept = median), color="blue")+
  theme_bw()+
  ylab('Density')+
  theme(aspect.ratio = 1)
```

```{r}
knitr::kable(mean_median_df)
```

# 7 

```{r}
ggplot(df, aes(x = age, y = charges))+
  geom_point(size=1)+
  theme_bw()+
  ggtitle('scatter plot: x = age, y = charges')+
  theme(aspect.ratio = 1, axis.text.x = element_text(size = 14),
        plot.title = element_text(hjust = 0.5))
```

# 8

```{r}
ggplot(df, aes(x = age, y = charges))+
  geom_point(size=1)+
  geom_smooth(method = lm) +
  theme_bw()+
  ggtitle('scatter plot: x = age, y = charges')+
  theme(aspect.ratio = 1, axis.text.x = element_text(size = 14),
        plot.title = element_text(hjust = 0.5))
```

# 9

```{r}
ggplot(df, aes(x = age, y = charges, col=smoker))+
    geom_point(size=1)+
    facet_wrap(.~smoker)+
    geom_smooth(method = lm, colour="black") +
    scale_color_brewer(palette = 'Set1')+
    theme_bw()+
    ggtitle('scatter plot: x = age, y = charges, facet_wrap = smoker')+
    theme(aspect.ratio = 1, axis.text.x = element_text(size = 14),
          plot.title = element_text(hjust = 0.5))
```

# 10

```{r}
ggplot(df, aes(x = bmi, y = charges, col=smoker))+
    geom_point(size=1)+
    facet_wrap(.~smoker)+
    geom_smooth(method = lm, colour="black") +
    scale_color_brewer(palette = 'Set1')+
    theme_bw()+
    ggtitle('scatter plot: x = bmi, y = charges, facet_wrap = smoker')+
    theme(aspect.ratio = 1, axis.text.x = element_text(size = 14),
          plot.title = element_text(hjust = 0.5))
```

# 11

How number of children is distributed across the age?

```{r}
p1 <- ggplot(df, aes(x = children, y = age, fill=children))+
  geom_boxplot()+
  theme_bw()+
  ggtitle('# children across age')+
  scale_fill_brewer(palette='Set1')+
  theme(aspect.ratio = 1, axis.text.x = element_text(size = 14),
        plot.title = element_text(hjust = 0.5))

p2 <- ggplot(df, aes(x = children, y = age, fill=children))+
  geom_violin()+
  theme_bw()+
  ggtitle('# children across age')+
  scale_fill_brewer(palette='Set1')+
  theme(aspect.ratio = 1, axis.text.x = element_text(size = 14),
        plot.title = element_text(hjust = 0.5))

ggarrange(p1, p2)
```

We can use boxplots and violin plots as well. Violin plots is better, since we can see the exact information about
target variable distribution. However, the perfect option is the combination of boxplot & violinplot at the same layout

```{r}
p2+geom_boxplot(width=0.1)
```

# 12

How number of children is distributed across females and males?


```{r}
ggplot(df, aes(children, ..count..))+
  geom_bar(aes(fill = sex), position = "dodge")+
  theme_bw()+
  ggtitle('# children across sex')+
  scale_fill_brewer(palette='Set1')+
  ylab('# of invidivuals')+
  theme(aspect.ratio = 1, axis.text.x = element_text(size = 14),
        plot.title = element_text(hjust = 0.5))
```

Barplots suitable for the visualization of categorical variables.

# 13

How smokers are distributed across regions (in terms of "raw" counts)?

```{r}
ggplot(df, aes(region, ..count..))+
  geom_bar(aes(fill = smoker), position = "dodge")+
  theme_bw()+
  ggtitle('# smokers across regions')+
  scale_fill_brewer(palette='Set1')+
  ylab('# of invidivuals')+
  theme(aspect.ratio = 1, axis.text.x = element_text(size = 14),
        plot.title = element_text(hjust = 0.5))
```

Since "raw counts" are comparible between different regions, we can inrepret this plot without switching
to the proportions (however, such type of questions should be answered using proportions, since number of observations
can vary between different groups).

Barplots suitable for the visualization of categorical variables.


# 14

```{r}
df <- df %>% 
  filter(age >= 21) %>% 
  mutate(age_group = ifelse(age <= 34, 'age: 21-34',
                            ifelse(age > 34 & age < 50, 'age: 35-49', 'age: 50+')))

ggplot(df, aes(x = bmi, y = log(charges)))+
  geom_point(alpha=0.4, color='#800080')+
  facet_wrap(.~age_group)+
  ylim(7, max(log(df$charges)))+
  geom_smooth(aes(group = age_group, color=age_group), method="lm")+
  theme_light()+
  theme(legend.position = 'bottom', plot.title = element_text(hjust = 0.5))+
  ggtitle('BMI ratio to the log(charges), data splitted by age groups')
```